# Results

<!-- PROBLEM ONE: Among all the carriers that provide commercial services in U.S., are there any carriers that have major delay problems? For example, the average delay time is over 60 minutes. -->


## Compare Carriers' On-Time Performance
As we can see from the figure below, `Delta Air Lines Inc.` have over 3,000 flights delayed during January 2019 from and to New York state, `American Airlines Inc.` also have more than 2,000 flights delayed, and the number is approximately 1.200 for `JetBlue Airways`. This is justifiable since these are the major airlines that conduct the most domestic flights, and the number of delayed flights, certainly, increases as the number of total conducted flights increases. Note that this figure only shows the number of times that delay occurs. On the other side, we need to evaluate the average severity of the delays, i.e. the average time of the delays for each carrier.



```{r, airline delay count}
library(ggplot2)
data1 = a1901_cleaning %>%
  filter(is.na(CARRIER_DELAY) == FALSE) %>%
  filter(ORIGIN_STATE_ABR == 'NY')
# data1 = filter(NY_Jan_cleaning,
#                is.na(NY_Jan_cleaning$CARRIER_DELAY) == FALSE)
  
delayCount <- data1 %>%
  count(data1$MKT_CARRIER_AIRLINE_ID, name = "count", sort = TRUE)
delayCount$CARRIER_AIRLINE_ID <- format(delayCount$`data1$MKT_CARRIER_AIRLINE_ID`)
delayCount <- delayCount %>% 
  mutate(CARRIER_AIRLINE_ID = factor(CARRIER_AIRLINE_ID, levels = rev(CARRIER_AIRLINE_ID[order(count)])))
delayCount <- delayCount[, -1]
ggplot(data = delayCount, mapping = aes(x = CARRIER_AIRLINE_ID, 
                                        y = count)) +
  geom_bar(stat = 'identity', fill = '#FFAFAF', color = '#5584AC') +
  labs(x = "Carrier Airline ID", y = 'Delay Count', title='Number of Delays in Jan. 2019 from and to NY')
```

<!-- The dataframe `NY_Jan_delayCount` shows the number of delays for each carrier within January. We plot a simple bar graph to manifest the result more explicitly. Note that this figure only shows the number of times that the delay occurs, on the other side, we need to evaluate the averaged severity of the delays, i.e. the averaged time of the delays for each carrier. -->



Therefore, we make another bar plot manifesting the average delay time for each carrier airline. Using the same dataset as above (flights from and to New York state during January 2019), we can see that `United Air Lines Inc` have the longest average delay time of 104.83 minutes. `Hawaiian Airlines Inc.`, `Spirit Air Lines`, `Delta Air Lines Inc.`, `JetBlue Airways`, `American Airlines Inc.` and `Frontier Airlines Inc.` all have average delay time longer than an hour.


```{r, airline delay average}
delayAverage <- data1
delayAverage$delayTime <- rowSums(delayAverage[c(25:29)]) #changed 27:31
delayAverage <- delayAverage %>%
  group_by(delayAverage$MKT_CARRIER_AIRLINE_ID) %>%
  summarise(delayMean = mean(delayTime))

delayAverage$CARRIER_AIRLINE_ID <- format(delayAverage$`delayAverage$MKT_CARRIER_AIRLINE_ID`)

delayAverage <- delayAverage %>% 
  mutate(CARRIER_AIRLINE_ID = factor(CARRIER_AIRLINE_ID, levels = rev(CARRIER_AIRLINE_ID[order(delayMean)])))
delayAverage <- delayAverage[, -1]

ggplot(data = delayAverage, mapping = aes(x = CARRIER_AIRLINE_ID, y = delayMean)) +
  geom_bar(stat = 'identity', fill = ifelse(delayAverage$delayMean >= 60, '#95D1CC', '#577BC1'), color = '#161853') +
  geom_text(aes(label=round(delayMean, 2)), position=position_dodge(width=0.9), vjust=-0.25) +
  labs(x = "Carrier Airline ID", y = 'Average Delay Time (min)', title='Average Delay Time for Each Carrier Airline')
```



Moreover, we hope to explore the results for concrete delay reasons, i.e. `CARRIER_DELAY`, `WEATHER_DELAY`, `NAS_DELAY`, `SECURITY_DELAY` and `LATE_AIRCRAFT_DELAY`. For each reason, we make two figures showing the count of delays and average delay time with respect to each carrier.  

From the figures, we can find that the most common and severe reason causing flight delays is `LATE_AIRCRAFT_DELAY`, which presumably can be due to the late arrival or preceding flights. There are almost 2,000 `Delta Air Lines Inc.` flights are delayed due to this reason, and the average waiting time exceeds 2 hours for `Hawaiian Airlines Inc.` flights.   

The second most common reason is `NAS_DELAY`, i.e., national air system delay, but the average time caused by this is typically not as long as that caused by other reasons, mostly within an hour.  

`CARRIER_DELAY` is the third most common reason that causes delays. This happens a lot of times for `Delta Air Lines Inc.`, `JetBlue Airways`, and `American Airlines Inc.`. On average, when this happens, the waiting time for `United Air Lines Inc.` flights is over 100 minutes. 

`WEATHER_DELAY` is less frequent than the three above. `Delta Air Lines Inc.` and `American Airlines Inc.` have about 150 delays caused by weather, which is a very small proportion of all delays. However, the time is highly variable- the longest is 118 minutes for `Delta Air Lines Inc.`, and the shortest is only 12.5 minutes.  

Among all 5 reasons, `SECURITY_DELAY` is the rarest, and only occurs 15 times. Therefore, we do not discuss this reason in detail.

```{r, delay reason}
library(dplyr)
library(tidyverse)
library(gridExtra)
library(patchwork)
delayDetails <- data1
# setwd("~/Desktop")
# changed 26 to 24!!
for (i in 1:5) {
  delayDetail = filter(delayDetails, delayDetails[,(i+24)] != 0)
  delayDetailCount <- delayDetail %>%
    count(delayDetail$MKT_CARRIER_AIRLINE_ID, name = "count", sort = TRUE)
  delayDetailCount$CARRIER_AIRLINE_ID <- format(delayDetailCount$`delayDetail$MKT_CARRIER_AIRLINE_ID`)
  delayDetailCount <- delayDetailCount %>% 
    mutate(CARRIER_AIRLINE_ID = factor(CARRIER_AIRLINE_ID, levels = rev(CARRIER_AIRLINE_ID[order(count)])))
  delayDetailCount <- delayDetailCount[, -1]
  p1 <- ggplot(data = delayDetailCount, 
               mapping = aes(x = CARRIER_AIRLINE_ID, y = count)) +
    geom_bar(stat = 'identity', fill = '#FFAFAF', color = '#5584AC') +
    labs(x = "ID", y = 'COUNT') + 
    ggtitle(paste(colnames(delayDetails)[i+24], 'COUNT'))
  
  delayDetailAverage <- delayDetail
  delayDetailAverage$delayDetailCopy <- delayDetailAverage[,(i+24)]
  delayDetailAverage <- delayDetailAverage %>%
    group_by(delayDetailAverage$MKT_CARRIER_AIRLINE_ID) %>%
    summarise(delayDetailMean = mean(delayDetailCopy))
  delayDetailAverage$CARRIER_AIRLINE_ID <- format(delayDetailAverage$`delayDetailAverage$MKT_CARRIER_AIRLINE_ID`)
  delayDetailAverage <- delayDetailAverage %>% 
    mutate(CARRIER_AIRLINE_ID = factor(CARRIER_AIRLINE_ID, levels = rev(CARRIER_AIRLINE_ID[order(delayDetailMean)])))
  delayDetailAverage <- delayDetailAverage[, -1]
  p2 <- ggplot(data = delayDetailAverage, 
               mapping = aes(x = CARRIER_AIRLINE_ID, 
                             y=delayDetailMean)) +
    geom_bar(stat = 'identity', 
             fill = ifelse(delayDetailAverage$delayDetailMean >= 60,
                           '#95D1CC', '#577BC1'), color = '#161853') +
    geom_text(aes(label=round(delayDetailMean, 2)), 
              position=position_dodge(width=0.6), vjust=-0.25) +

    labs(x = "ID", y = 'AVE_TIME') + 
    ggtitle(paste(colnames(delayDetails)[i+24],'AVE_TIME'))
  
  p1 | p2
  p <- p1 + p2 + plot_layout(ncol = 2)
  print(p)
  
  # if (i == 1) {
  #   p <- p1 + p2
  # }
  # else{
  #   p <- p + p1 + p2  #很丑 #(很可爱了哈哈
  # }
  
  # filename <- paste(colnames(delayDetails)[(26+i)], ".pdf", sep = "")
  # ggsave(filename = filename)
  
}
```



<!-- PROBLEM TWO: Are early flights or late flights more likely to delay? Do origin and destination influence the on-time performance? -->

## On-Time Performance Accross the Day

In addition to comparing the differences among the various airlines, we also compared the on-time performance of flights at difference time of a day. We divide the 24 hours of a day into 8 groups, one group every three hours, and calculated the average delay time of departure and arrival. As we can see in the plot below, flight performance of 2020 is obviously better than in 2019. The green horizontal line at level 0 indicates the flight delays or not. Specifically, flights in 2020 almost arrive in advance to their destination no matter they departs lately or not. Unfortunately, flights in 2019 almost depart lately. Moreover, late flights seem to performance worse than early flights in general.



```{r, performance based on CRS_DEP_TIME}
# flight performance per 3h
# want to see if early or late flights delay more
# divide into groups and see delay performance?
# use NY_2019_cleaning


time_lv <- c("12am~3am", "3am~6am", "6am~9am", "9am~12pm",
             "12pm~3pm", "3pm~6pm", "6pm~9pm", "9pm~12am")

from_NY_2019 <- US_2019_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY') ) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr=cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300))) %>% 
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean))
levels(from_NY_2019$dep_gr) = time_lv

from_NY_2020 <- US_2020_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY')) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr=cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300))) %>% 
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean))
levels(from_NY_2020$dep_gr) = time_lv


# ggplot(from_NY_2019, aes(x=dep_gr)) + 
#   geom_point(aes(y=DEP_DELAY_Mean, color='Departure')) + 
#   geom_point(aes(y=ARR_DELAY_Mean, color='Arrival')) +
#   scale_colour_manual("", breaks=c('Departure','Arrival'), values=c('red','blue')) +
#   labs(x="CRS Departure Time", y="Average Delay (min)", title='Flight Performance from NY in 2019')
# 
# ggplot(from_NY_2020, aes(x=dep_gr)) + 
#   geom_point(aes(y=DEP_DELAY_Mean, color='Departure')) + 
#   geom_point(aes(y=ARR_DELAY_Mean, color='Arrival')) +
#   scale_colour_manual("", breaks=c('Departure','Arrival'), values=c('red','blue')) +
#   labs(x="CRS Departure Time", y="Average Delay (min)", title='Flight Performance from NY in 2020')

from_NY_2019$year <- "2019"
from_NY_2020$year <- "2020"
from_NY <- rbind(from_NY_2019, from_NY_2020)

ggplot(from_NY, aes(x=dep_gr, shape = year)) + 
  geom_point(aes(y = DEP_DELAY_Mean, color = 'Departure', size = 1)) + 
  geom_point(aes(y = ARR_DELAY_Mean, color = 'Arrival', size = 1)) +
  geom_hline(yintercept = 0, color = "#116530") + 
  scale_colour_manual("", breaks=c('Departure','Arrival'), values=c('#DD4A48','#064663')) +
  scale_discrete_manual(values = c(4,18), aesthetics = 'shape') +
  labs(x="CRS Departure Time", y="Average Delay (min)", title='Early / Late Flight Performance from NY') +
  scale_size_continuous(guide = "none")
```

Next, we explore the difference among flights to different destinations. We divide them into groups by state in the following graph. The trend in the previous graph is still easy to see in this one- flights that depart later in a day tend to have longer average delay time, and that the on-time performance for flights in 2020 is significantly better than that in 2019. Note that flights to some destinations e.g. HI, MT, NM, are only provided in certain periods, which may caused by the distance and passenger volume.


```{r}
time_lv <- c("12am~3am", "3am~6am", "6am~9am", "9am~12pm",
             "12pm~3pm", "3pm~6pm", "6pm~9pm", "9pm~12am")

time_lv_index <- c("0", "1", "2", "3", "4","5", "6", "7")

from_NY_2019_dest <- US_2019_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY') ) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr=cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), DEST_STATE_ABR) %>% 
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean))
levels(from_NY_2019_dest$dep_gr) = time_lv_index

from_NY_2020_dest <- US_2020_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY')) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr=cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), DEST_STATE_ABR) %>% 
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean))
levels(from_NY_2020_dest$dep_gr) = time_lv_index

from_NY_2019_dest$year <- "2019"
from_NY_2020_dest$year <- "2020"
from_NY_dest <- rbind(from_NY_2019_dest, from_NY_2020_dest)

ggplot(from_NY_dest, aes(x=dep_gr)) + 
  geom_point(aes(y = DEP_DELAY_Mean, color = year)) + 
  geom_hline(yintercept = 0, color = "#116530") + 
  scale_colour_manual(values=c('#DD4A48','#FBD148')) +
  labs(x="CRS Departure Time", y="Average Delay (min)", title='Early / Late Flight Performance from NY to US') +
  facet_wrap(~DEST_STATE_ABR, ncol=10)
  
```






```{r}
library(stringr)
from_NY_2019_month <- US_2019_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY') ) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr = cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), mon = str_sub(FL_DATE,6,7)) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean))
levels(from_NY_2019_month$dep_gr) = time_lv_index

from_NY_2020_month <- US_2020_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY')) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr = cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), mon = str_sub(FL_DATE,6,7)) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean))
levels(from_NY_2020_month$dep_gr) = time_lv_index

from_NY_2019_month$year <- "2019"
from_NY_2020_month$year <- "2020"
from_NY_month <- rbind(from_NY_2019_month, from_NY_2020_month)

from_NY_month <- within(from_NY_month, mon <- factor(mon, levels = c("12", "03", "06", "09","01", "04", "07", "10","02", "05", "08", "11")))
with(from_NY_month, levels(mon))

ggplot(from_NY_month, aes(x=dep_gr)) + 
  geom_point(aes(y = DEP_DELAY_Mean, color = year)) + 
  geom_hline(yintercept = 0, color = "#116530") + 
  scale_colour_manual(values=c('#DD4A48','#FBD148')) +
  labs(x="CRS Departure Time", y="Average Delay (min)", title='Monthly Early / Late Flight Performance from NY') +
  facet_wrap(~mon, ncol=4)

```

The plot above shows flights performance in each months, which also yields the two conclusions mentioned before. Moreover, some other interesting observations are: (1) in winter (Dec., Jan., Feb.), flight performance is in general worse than in other seasons, especially for those late flights, which may due to the cold weather. (2) Fall (Sept., Oct., Nov.) seems to have the best flight performance.

```{r}
library(epade)
library(reshape2)

from_NY_2019_allu <- US_2019_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY') ) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr = cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), FL_DATE) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean)) %>%
  group_by(mon = str_sub(FL_DATE,6,7))
levels(from_NY_2019_allu$dep_gr) = time_lv


from_NY_2020_allu <- US_2020_cleaning %>%
  filter((ORIGIN_CITY_NAME=='New York, NY')) %>%
  arrange(CRS_DEP_TIME) %>%
  group_by(dep_gr = cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), FL_DATE) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean)) %>%
  group_by(mon = str_sub(FL_DATE,6,7))
levels(from_NY_2020_allu$dep_gr) = time_lv

from_NY_2019_allu$YEAR <- "2019"
from_NY_2020_allu$YEAR <- "2020"
from_NY_2019_allu$DATE <- str_sub(from_NY_2019_allu$FL_DATE,9,10)
from_NY_2020_allu$DATE <- str_sub(from_NY_2020_allu$FL_DATE,9,10)


from_NY_2019_dat = from_NY_2019_allu[, c(1,3,5)]
from_NY_2019_dat <- from_NY_2019_dat[, c("mon","dep_gr","DEP_DELAY_Mean")]
names(from_NY_2019_dat)[names(from_NY_2019_dat) == 'DEP_DELAY_Mean'] <- 'value'
par(las=2,cex=1.5)
tmp = dcast(from_NY_2019_dat, mon~dep_gr, mean)
tmp_data = as.matrix(tmp[, -1])
rownames(tmp_data) = tmp[, 1]
p1 <- bar3d.ade(t(tmp_data), wall = 1, xw = 0.3, zw = 1.2, xlab = "month", ylab = "ave-delay", main = "2019 month-period averaged delay")

from_NY_2020_dat = from_NY_2020_allu[, c(1,3,5)]
from_NY_2020_dat <- from_NY_2020_dat[, c("mon","dep_gr","DEP_DELAY_Mean")]
names(from_NY_2020_dat)[names(from_NY_2020_dat) == 'DEP_DELAY_Mean'] <- 'value'
par(las=2,cex=1.5)
tmp2 = dcast(from_NY_2020_dat, mon~dep_gr, mean)
tmp_data2 = as.matrix(tmp2[, -1])
rownames(tmp_data2) = tmp2[, 1]
p2 <- bar3d.ade(t(tmp_data2), wall = 3, xw = 0.6, zw = 1.8, xlab = "month", ylab = "ave-delay", main = "2020 month-period averaged delay")

```

3D plot is more powerful to show the variation of a quantity grouped by multiple dimensions. Flights in June, July and August from 9pm to 12 pm have the longest average delay time.

## Flights to Three Airports near NYC (JFK, LGA, EWR)
<!-- PROBLEM THREE: Specifically for NYC. If someone wants to come to NYC, which airport should s/he choose to fly to based on the origin city? -->

Columbia University is located in the City of New York, and there are three major airports near the city, which are John F. Kennedy International Airport (JFK), LaGuardia Airport (LGA) and Newark Liberty International Airport (EWR). Suppose a person wants to travel to New York, how should he/she choose which airport to fliy to? We make a graph to illustrate the differnce, facet by each state that has flight towards the three airports.


```{r, to NYC problem}
to_NYC_2019 <- US_2019_cleaning %>% 
  filter((DEST=='LGA') | (DEST=='JFK') | (DEST=='EWR')) %>%
  arrange(ORIGIN_STATE_ABR, CRS_DEP_TIME) %>%
  # group_by(dep_gr = cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), ORIGIN_STATE_ABR, DEST) %>%
  group_by(ORIGIN_STATE_ABR, DEST) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean)) %>%
  # arrange(dep_gr, ORIGIN_STATE_ABR)
  arrange(ORIGIN_STATE_ABR)
# levels(to_NYC_2019$dep_gr) = time_lv_index


to_NYC_2020 <- US_2020_cleaning %>% 
  filter((DEST=='LGA') | (DEST=='JFK') | (DEST=='EWR')) %>%
  group_by(ORIGIN_STATE_ABR, DEST) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean)) %>%
  arrange(ORIGIN_STATE_ABR)

to_NYC_2019$year <- "2019"
to_NYC_2020$year <- "2020"

to_NYC <- rbind(to_NYC_2019, to_NYC_2020)


ggplot(to_NYC, aes(x=DEST)) + 
  geom_point(aes(y = DEP_DELAY_Mean, color = year)) + 
  geom_hline(yintercept = 0, color = "#116530") + 
  scale_colour_manual(values=c('#DD4A48','#FBD148')) +
  labs(x="Destination Airport", y="Average Departure Delay (min)", title='Flight Performance to NYC') +
  facet_wrap(~ORIGIN_STATE_ABR, ncol=10)



# ggplot(to_NYC_2019, aes(x=DEST)) + 
#   geom_point(aes(y = DEP_DELAY_Mean, color = DEST)) + 
#   geom_hline(yintercept = 0, color = "#116530") + 
#   # scale_colour_manual(values=c('#DD4A48','#FBD148')) +
#   labs(x="CRS Departure Time", y="Average Departure Delay (min)", title='Flight Performance to NYC') +
#   # facet_wrap(~DEST, ncol=4) +
#   facet_wrap(~ORIGIN_STATE_ABR, ncol=10)



# to_NYC_2020 <- US_2020_cleaning %>% 
#   filter((DEST=='LGA') | (DEST=='JFK')| (DEST=='EWR')) %>%
#   arrange(ORIGIN_STATE_ABR, CRS_DEP_TIME) %>%
#   group_by(dep_gr = cut(CRS_DEP_TIME, breaks=seq(0, 2400, by=300)), ORIGIN_STATE_ABR, DEST) %>%
#   summarise_at(vars(DEP_DELAY, ARR_DELAY), list(Mean=mean)) %>%
#   arrange(dep_gr, ORIGIN_STATE_ABR)
# levels(to_NYC_2020$dep_gr) = time_lv_index

# ggplot(to_NYC_2019, aes(x=DEST)) + 
#   geom_point(aes(y = DEP_DELAY_Mean, color=ORIGIN_STATE_ABR)) + 
#   geom_hline(yintercept = 0, color = "#116530") + 
#   labs(x="Airpot", y="Average Delay (min)", 
#        title='Flight Performance from US to NYC')
```

```{r}
# which airline is the fastest?
# DISTANCE / ACTUAL_ELAPSED_TIME
# or AIR_TIME

airline_19 <- US_2019_cleaning %>%
  group_by(MKT_CARRIER_AIRLINE_ID) %>%
  mutate(speed = DISTANCE / AIR_TIME *60) %>%
  summarise_at(vars(DEP_DELAY, ARR_DELAY, speed), 
               list(Mean=mean))

ggplot(airline_19) + 
  geom_point(aes(x=MKT_CARRIER_AIRLINE_ID, y=speed_Mean)) +
  labs(x="Carrier ID", y="Average Speed (mph)", 
       title = "Average Speed of US Domestic Flights")

```














